---
title: "iHg - Markov chain Monte Carlo Calibration"
author: "Nan-Hung Hsieh" modified by "Yu-Sheng Lin"
date: "2022/02/05 (update: `r Sys.Date()`)"
# Run Monte Carlo and Uncertainty analysis for better estimation the range for parameter values
# if needed (the last section)
# cd ./mcsim-6.2.0/IRIS_iHg/modeling/MiceHier


output: 
  html_document:
  fig_caption: yes  
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#setwd('..')
#wd <- getwd()
#knitr::opts_knit$set(root.dir =  wd)
```

# 0 Prerequisites
```{r, include=FALSE}
source("MCSim/function.R")
getwd()
```

# Import the R packages
```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(bayesplot)
library(corrplot)
library(sensitivity)
library(pksensi)
library(data.table)
library(coda)
theme_set(theme_bw())
```

##### 2. QA/QC check


##### 2.1 Balance check and gavage calculation check --- dynamic body weight change for Mice
```{r}
model <- "iHgMiceBW.model"
input <- "iHgMiceNtp1993BW.in"
RMCSim::makemcsim(model, dir = "modeling")
test <- mcsim(model = model, input = input, dir = "modeling")
testex <- readsims(test,  exp=1) 
plot (testex$Time, testex$BW, xlim = c(0, 24), ylim = c(0.0001, 0.5),  type="l", lwd=2, xlab="Time", ylab="Body weight", log='y')
```

###### 3.1 Model verification of Individual studies ########

#################### Mice study #########################

##### 1. Female NMRI w/ single IV dose of 400 µg Hg/kg/d (1982-Aaseth):    
```{r}
Aaseth <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceAasethBW.in", dir = "modeling")
Aaseth <- readsims(Aaseth, exp=1) 
```

##### 1. Female NMRI w/ single IV dose of 400 µg Hg/kg/d (1982-Aaseth): Kidney    
```{r}
data_t <- c(72,	96);
data_c <- c(3.319728,	2.560555);

plot(data_t, data_c, main = "Mice, 1982 Aaseth, IV=400", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 288), ylim = c(0.1, 20), log='y')
lines(Aaseth$Time, Aaseth$CKU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 1. Female NMRI w/ single IV dose of 400 µg Hg/kg/d (1982-Aaseth): brain  
```{r}
data_t <- c(72,	96);
data_c <- c(0.13055565,0.03);

plot(data_t, data_c, main = "Mice, 1982 Aaseth, IV=400", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 288), ylim = c(0.001, 10), log='y')
lines(Aaseth$Time, Aaseth$CBrnU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 1. Female NMRI w/ single IV dose of 400 µg Hg/kg/d (1982-Aaseth): urine
```{r}
data_t <- c(24 , 48 , 72 , 96);
data_c <- c(1.60056272 , 2.220332931 , 2.602229033 , 2.772038576);

plot(data_t, data_c, main = "Mice, 1982 Aaseth, IV=400", xlab = "Time (hr)", ylab = "Urine (ug)", xlim = c(0, 100), ylim = c(0.01, 10), log='y')
lines(Aaseth$Time, Aaseth$Aurine)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 2. Male ICR w/ single oral dose of 1000 µg Hg/kg/d; Matrix: blood (Nielsen, 1992, 5925484)    
```{r}
NielsenD <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceNielsenBW.in", dir = "modeling")
NielsenD <- readsims(NielsenD, exp=1) 
```

#2# Blood (Nielsen, 1992, 5925484)   
```{r}
data_t <- c(24 , 72 , 240 , 480);
data_c <- c(0.014794515 , 0.005133886 , 0.001263642 , 0.000373918);

plot(data_t, data_c, main = "Nielsen 92, Mice, Single oral of 1000 ug/kg", xlab = "Time (hr)", ylab = "CBld (ug/g)", xlim = c(0, 480), ylim = c(0.0001, 1), log='y')
lines(NielsenD$Time, NielsenD$CBldU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

#2# Brain (Nielsen, 1992, 5925484)   
```{r}
data_t <- c(24 , 48, 72 , 240 , 720);
data_c <- c(0.083859475,	0.017162917,	0.047963081,	0.11661534,	0.012529468);

plot(data_t, data_c, main = "Nielsen 92, Mice, Single oral of 1000 ug/kg", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 720), ylim = c(0.000001, 1), log='y')
lines(NielsenD$Time, NielsenD$CBrnU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 3000-75000 µg Hg/kg/d for 7 weeks (Dieter, 1983):  
```{r}
Dieter <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceDieterBW.in", dir = "modeling")
Dieter3000 <- readsims(Dieter, exp=1)
Dieter15000 <- readsims(Dieter, exp=2)
Dieter75000 <- readsims(Dieter, exp=3)
```

##### 3. Male B6C3Fl mice with oral gavage of 3000 µg Hg/kg/d for 7 weeks (Dieter, 1983): blood 
```{r}
data_t <- c(168, 504, 1176);
data_c <- c(0.1, 0.1, 0.4);
sd <- c(0, 0, 0.3)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 3000, male mice", xlab = "Time (hr)", ylab = "CBld (ug/g)", xlim = c(0, 1200), ylim = c(0.01, 10), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter3000$Time, Dieter3000$CBldU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 3000 µg Hg/kg/d for 7 weeks (Dieter, 1983): kidney 
```{r}
data_t <- c(168, 504, 840, 1176);
data_c <- c(3.9, 3.1, 1, 4);
sd <- c(1.3, 0.7, 0.1, 0.5)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 3000, male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 1200), ylim = c(0.01, 100), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter3000$Time, Dieter3000$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 15000 µg Hg/kg/d for 7 weeks (Dieter, 1983): blood 
```{r}
data_t <- c(168, 504, 840, 1176);
data_c <- c(0.2, 0.5, 0.7, 0.6);
sd <- c(0.2, 0.3, 0.2, 0.3)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 15000, male mice", xlab = "Time (hr)", ylab = "CBld (ug/g)", xlim = c(0, 1200), ylim = c(0.01, 20), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter15000$Time, Dieter15000$CBldU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 15000 µg Hg/kg/d for 7 weeks (Dieter, 1983): kidney 
```{r}
data_t <- c(168, 504, 840, 1176);
data_c <- c(14.9, 16.9, 22.7, 20.7);
sd <- c(2, 2, 9, 4)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 15000, male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 1500), ylim = c(0.01, 1000), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter15000$Time, Dieter15000$CKU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 75000 µg Hg/kg/d for 7 weeks (Dieter, 1983): blood 
```{r}
data_t <- c(168, 504, 840, 1176);
data_c <- c(0.3, 0.5, 1.1, 1.4);
sd <- c(0.1, 0.1, 0.2, 0.4)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 75000, male mice", xlab = "Time (hr)", ylab = "CBld (ug/g)", xlim = c(0, 1200), ylim = c(0.01, 100), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter75000$Time, Dieter75000$CBldU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Male B6C3Fl mice with oral gavage of 75000 µg Hg/kg/d for 7 weeks (Dieter, 1983): kidney 
```{r}
data_t <- c(168, 504, 840, 1176);
data_c <- c(74.9, 68.9, 57.7, 57.7);
sd <- c(9,	11,	8,	8)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Dieter, 1983, Oral = 75000, male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 1200), ylim = c(0.01, 1000), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Dieter75000$Time, Dieter75000$CKU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male and Female B6C3Fl mice with oral gavage of 20000 to 1250 µg HgCl2 /kg/d for 6 months (Ntp 1993):   
```{r}
Ntp1993 <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceNtp1993BW.in", dir = "modeling")
Ntp1993Male1250  <- readsims(Ntp1993, exp=1)
Ntp1993Male5000  <- readsims(Ntp1993, exp=2)
Ntp1993Male20000 <- readsims(Ntp1993, exp=3)
Ntp1993Female1250  <- readsims(Ntp1993, exp=4)
Ntp1993Female5000  <- readsims(Ntp1993, exp=5)
Ntp1993Female20000 <- readsims(Ntp1993, exp=6)
```

##### 4. Male B6C3Fl mice with oral gavage of 1250 µg HgCl2 (or 925 ug Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(1.2,	0.93,	0.8);
sd <- c(1.22, 0.25, 0.09)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 1250, Male mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 10), log='y')
lines(Ntp1993Male1250$Time, Ntp1993Male1250$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 1250 µg HgCl2 (or 925 ug Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(6.98,	7.02,	6.97);
sd <- c(2.08, 1.60, 2.01)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 1250, Male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(1, 50), log='y')
lines(Ntp1993Male1250$Time, Ntp1993Male1250$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 5000 µg HgCl2 (or 3695 ug Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(2.66,	9.71,	2.93);
sd <- c(0.46, 3.52, 0.84)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 5000, Male mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 50), log='y')
lines(Ntp1993Male5000$Time, Ntp1993Male5000$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 5000 µg HgCl2 (or 3695 ug Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(36.43,	26.43,	35.68);
sd <- c(10.5, 7.75, 6.84)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 5000, Male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 200), log='y')
lines(Ntp1993Male5000$Time, Ntp1993Male5000$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(9.88,	9.71,	10.55);
sd <- c(1.81, 3.52, 3.94)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Male mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(1, 100), log='y')
lines(Ntp1993Male20000$Time, Ntp1993Male20000$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(111.73,	104.73,	86.88);
sd <- c(12.9, 10.1, 9.83)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(1, 500), log='y')
lines(Ntp1993Male20000$Time, Ntp1993Male20000$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Male B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): brain 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(0.24,	0.35,	0.19);
sd <- c(0.06, 0.11, 0.10)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Male mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 20), log='y')
lines(Ntp1993Male20000$Time, Ntp1993Male20000$CBrnU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 1250 µg HgCl2 (or 925 ug Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(0.93,	0.84,	1.04);
data_c_Ntp1993Female1250liver <- data_c
sd <- c(0.54, 0.55, 0.25)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 1250, Female mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 10), log='y')
lines(Ntp1993Female1250$Time, Ntp1993Female1250$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 1250 µg HgCl2 (or 925 ug Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(7.29,	8.32,	9.86);
data_c_Ntp1993Female1250kidney <- data_c
sd <- c(0.93, 0.86, 1.78)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 1250, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(1, 50), log='y')
lines(Ntp1993Female1250$Time, Ntp1993Female1250$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 5000 µg HgCl2 (or 3695 ug Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(2.74,	3.82,	3.3);
data_c_Ntp1993Female5000liver <- data_c
sd <- c(0.55, 0.82, 0.31)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 5000, Female mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 50), log='y')
lines(Ntp1993Female5000$Time, Ntp1993Female5000$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 5000 µg HgCl2 (or 3695 ug Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(26.64,	23.49,	40.16);
data_c_Ntp1993Female5000kidney <- data_c
sd <- c(4.77, 7.51, 7.75)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 5000, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 200), log='y')
lines(Ntp1993Female5000$Time, Ntp1993Female5000$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): liver 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(8.7,	13.46,	13.32);
data_c_Ntp1993Female20000liver <- data_c
sd <- c(1.67, 4, 2.73)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Female mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 4320), ylim = c(1, 100), log='y')
lines(Ntp1993Female20000$Time, Ntp1993Female20000$CLU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): kidney 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(87.94,	97.19,	87.96);
data_c_Ntp1993Female20000kidney <- data_c
sd <- c(18.2, 14.4, 9.68)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 4320), ylim = c(1, 500), log='y')
lines(Ntp1993Female20000$Time, Ntp1993Female20000$CKU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Female B6C3Fl mice with oral gavage of 20000 µg HgCl2 (or 14775 µg Hg)/kg/d for 6 months (Ntp 1993): brain 
```{r}
data_t <- c(1428, 2868,	4308);
data_c <- c(0.30,	0.49,	0.81);
data_c_Ntp1993Female20000brain <- data_c
sd <- c(0.15, 0.14, 0.21)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 6 month, Oral = 20000, Female mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 4320), ylim = c(0.1, 20), log='y')
lines(Ntp1993Female20000$Time, Ntp1993Female20000$CBrnU)
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Male and Female B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993):  
```{r}
Ntp1993wks <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceNtp1993wksBW.in", dir = "modeling")
Ntp1993Male29550 <- readsims(Ntp1993wks, exp=1)
Ntp1993Female29550 <- readsims(Ntp1993wks, exp=2)
```

##### 5. Male B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): liver 
```{r}
data_t <- c(384)
data_c <- c(34.35)
data_c_Ntp1993Male29550liver <- data_c
sd <- c(21.57)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Male mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 400), ylim = c(1, 400), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Male29550$Time, Ntp1993Male29550$CLU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Male B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): kidney 
```{r}
data_t <- c(384);
data_c <- c(170.892);
data_c_Ntp1993Male29550kidney <- data_c
sd <- c(47.7)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 400), ylim = c(1, 500), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Male29550$Time, Ntp1993Male29550$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Male B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): brain 
```{r}
data_t <- c(384);
data_c <- c(0.47);
data_c_Ntp1993Male29550brain <- data_c
sd <- c(0.376)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Male mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 400), ylim = c(0.1, 20), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Male29550$Time, Ntp1993Male29550$CBrnU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Female B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): liver 
```{r}
data_t <- c(384);
data_c <- c(29.18);
data_c_Ntp1993Female29550liver <- data_c
sd <- c(10.05)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Female mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 400), ylim = c(1, 200), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Female29550$Time, Ntp1993Female29550$CLU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Female B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): kidney 
```{r}
data_t <- c(384);
data_c <- c(115.592);
data_c_Ntp1993Female29550kidney <- data_c
sd <- c(8.7)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 400), ylim = c(1, 500), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Female29550$Time, Ntp1993Female29550$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 5. Female B6C3Fl mice with oral gavage of 29550 µg Hg (as 40 mg HgCl2) /kg/d for 16 days (Ntp 1993): brain 
```{r}
data_t <- c(384);
data_c <- c(0.493);
data_c_Ntp1993Female29550brain <- data_c
sd <- c(0.224)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Ntp 1993, 16 day, Oral = 29550, Female mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 400), ylim = c(0.1, 20), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Ntp1993Female29550$Time, Ntp1993Female29550$CBrnU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 6. Male C57BL/6J mice oral gavage of 10000 µg HgCl2 (or 7390 ug Hg)/kg/d for 10 days (Wang 2013):  
```{r}
Wang10000 <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceWang2013BW.in", dir = "modeling")
Wang <- readsims(Wang10000, exp=1)
```

##### 6. Male C57BL/6J mice oral gavage of 10000 µg HgCl2 (or 7390 ug Hg)/kg/d for 10 days (Wang 2013):
# liver 
```{r}
data_t <- c(240);
data_c <- c(5.560);
data_c_Wang10000liver <- data_c
sd <- c(4.195)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Wang 2013, 10 day, Oral = 10000, Male mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 300), ylim = c(1, 100), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Wang10000$Time, Wang10000$CLU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 6. Male C57BL/6J mice oral gavage of 10000 µg HgCl2 (or 7390 ug Hg)/kg/d for 10 days (Wang 2013):
# kidney
```{r}
data_t <- c(240);
data_c <- c(23.578);
data_c_Wang10000kidney <- data_c
sd <- c(1.54499)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Wang 2013, 10 day, Oral = 10000, Male mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 300), ylim = c(1, 1000), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Wang10000$Time, Wang10000$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 6. Male C57BL/6J mice oral gavage of 10000 µg HgCl2 (or 7390 ug Hg)/kg/d for 10 days (Wang 2013):
# brain
```{r}
data_t <- c(240);
data_c <- c(0.189);
data_c_Wang10000brain <- data_c
sd <- c(0.04)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Wang 2013, 10 day, Oral = 10000, Male mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 300), ylim = c(0.01, 10), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Wang10000$Time, Wang10000$CBrnU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 6. Male C57BL/6J mice oral gavage of 10000 µg HgCl2 (or 7390 ug Hg)/kg/d for 10 days (Wang 2013):
# serum/blood
```{r}
data_t <- c(240);
data_c <- c(0.400);
data_c_Wang10000blood <- data_c
sd <- c(0.104)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Wang 2013, 10 day, Oral = 10000, Male mice", xlab = "Time (hr)", ylab = "CBlood (ug/g)", xlim = c(0, 300), ylim = c(0.01, 10), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Wang10000$Time, Wang10000$CBldU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


##### 7. Female A.SW (H-2s) mice drinking water of 148 µg Hg/kg/d for 2 days 
## (Havarinasab, 2007 and Hultman, 1998):  
```{r}
Havar148 <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceHultmanBW.in", dir = "modeling")
Havar <- readsims(Havar148, exp=1)
```

##### 7. Female A.SW (H-2s) mice with 48 hrs, 148 ug/kg (Havarinasab, 2007)
# kidney
```{r}
data_t <- c(24, 36, 48);
data_c <- c(0.5565217391,	2.12173913,	2.504347826);

sd <- c(0.5, 0.5, 0.5)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Havarinasab, 2007, 48 hrs, Oral = 148, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 50), ylim = c(0.0001, 30), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Havar$Time, Havar$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


##### 8. Female Swiss Albino  mice drinking water of 6000 µg Hg/kg/d for 2 days 
## (Sin 1990):  
```{r}
Sin6000 <- mcsim(model = "iHgMiceBW.model", input = "iHgMiceSin1990BW.in", dir = "modeling")
Sin<- readsims(Sin6000, exp=1)
```

##### 8. Female Swiss Albino  mice with 10 days, 6000 ug Hg/kg (Sin 1990)
# kidney
```{r}
data_t <- c(240);
data_c <- c(57.82);
data_c_Sin6000kidney <- data_c
sd <- c(19.159)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Sin 1990, 10 days, Oral = 6000, Female mice", xlab = "Time (hr)", ylab = "CK (ug/g)", xlim = c(0, 240), ylim = c(0.01, 500), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Sin$Time, Sin$CKU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


##### 8. Female Swiss Albino  mice with 48 hrs, 6000 ug Hg/kg (Sin 1990)
# liver
```{r}
data_t <- c(240);
data_c <- c(8.4);
data_c_Sin6000liver <- data_c
sd <- c(1.715)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Sin 1990, 10 days, Oral = 6000, Female mice", xlab = "Time (hr)", ylab = "CL (ug/g)", xlim = c(0, 240), ylim = c(0.1, 100), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Sin$Time, Sin$CLU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


##### 8. Female Swiss Albino  mice with 48 hrs, 6000 ug Hg/kg (Sin 1990)
# brain
```{r}
data_t <- c(240);
data_c <- c(0.43);
data_c_Sin6000brain <- data_c
sd <- c(0.147)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Sin 1990, 10 days, Oral = 6000, Female mice", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 240), ylim = c(0.1, 100), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Sin$Time, Sin$CBrnU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


#### 2.5 Multi chains testing ### 
#### MCMC Modeling ###
### Using Ubuntu instead ###

# Single chain check
# Checking model fit using one single chain #
# Global evaluation/ "MCMC.check.out" has "Data" and "Predicion"
```{r}
single_1 <- fread("outputs/iHgMice_3365.out") |> as.data.frame()
tail(single_1)
```

# Single chain check
# Checking model fit using one single chain #
# Global evaluation/ "MCMC.check.out" has "Data" and "Predicion"

## Result diagnosis ##
# Trace plot
```{r}
str <- ceiling(nrow(single_1)/2) + 1
end <- nrow(single_1)
j <- c(str:end) # discard burn-in
par(mfrow = c(3,4), mar = c(2,2,2,1))
for (i in 2:63){
  plot(single_1[j,1], single_1[j,i], type = "l", main = names(single_1)[i])
}
```

# Auto correlation
```{r}
par(mfrow = c(3,4), mar = c(2,2,4,1))
for (i in 2:63){
  acf(single_1[j,i], main = names(single_1)[i])
  abline(0.5, 0, col=2, lty=2)
}
```

# Density plot
```{r}
par(mfrow = c(3,4), mar = c(2,2,2,1))
for (i in 2:63){
  single_1[j,i] %>% density() %>% plot(type = "l", main = names(single_1)[i])
}
```

# Correlation matrix
```{r}
single_1[,7:15] %>% cor() %>% corrplot(method = "number")
```

# Global evaluation/ "MCMC.check.out" has "Data" and "Predicion"
# Check simulation result with the parameter from the last iteration
```{r}
chk_1 <- read.delim("outputs/iHgMice_check_4880.out")
chk_1 <- chk_1[which(chk_1$Data>0),]
tail(chk_1)
```

# Overall fitting evaluation
```{r}
R_square <- cor(chk_1$Data, chk_1$Prediction)^2
plot(chk_1$Data, chk_1$Prediction, log = "xy",
     xlim = c(0.001, 1000), ylim = c(0.001, 1000),
     xlab = "Observation", ylab = "Prediction",
     main = paste("R2 =", round(R_square, digits = 3)))
abline(0,1)
```

# Individual measurement fitting evaluation 
# (1) amount of iHg in urine
```{r}
chk_1 %>% filter(Output_Var == "Aurine") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + ylim(1, 10) +
  labs(x = "Time (min)", y = "Urinary iHg (ug Hg)", title = "Accumulative iHg excreted into urine in mice")
```


# (3) concentratio of iHg in blood (ug/ml)
```{r}
chk_1 %>% filter(Output_Var == "CBldU") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() +  
  labs(x = "Time (min)", y = "Blood iHg (ug Hg/ml)", title = "Blood iHg in mice")
```

# (4) concentratio of iHg in kidney (ug/ml)
```{r}
chk_1 %>% filter(Output_Var == "CKU") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Kiney IHg (ug Hg/g)", title = "Kidney iHg in mice" )
```

# (5) concentratio of iHg in liver (ug/ml)
```{r}
chk_1 %>% filter(Output_Var == "CLU") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() +  ylim(0.1, 30) +
  labs(x = "Time (min)", y = "Liver IHg (ug Hg/g)", title = "Liver iHg in mice")
```

# (6) concentratio of iHg in brain (ug/ml)
```{r}
chk_1 %>% filter(Output_Var == "CBrnU") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() +    ylim(0.0001, 10) +
  labs(x = "Time (min)", y = "Brain IHg (ug/ml)", title = "Brain iHg in mice")
```

#### 2.5 Multi chains testing ### 
#### MCMC Modeling ###
### Using Ubuntu instead ###

```{r, eval=F}
Micejob1 <- fread("outputs/iHgMice_4880.out") |> as.data.frame()
Micejob2 <- fread("outputs/iHgMice_3365.out") |> as.data.frame()
Micejob3 <- fread("outputs/iHgMice_5916.out") |> as.data.frame()
Micejob4 <- fread("outputs/iHgMice_6734.out") |> as.data.frame()
x <- mcmc_array(data = list(Micejob1, Micejob2, Micejob3, Micejob4))
str <- ceiling(nrow(x)/2) + 1
end <- nrow(x)
j <- c(str:end) # discard burn-in
sample_x <- seq(50000, 100000, 10) + 1
x <- x[sample_x,,]
```

# Size of output summary
```{r}
dim(x)
mcmc_mice <- x
save(mcmc_mice, file = "mcmc_mice.RData")
```

# traceplot and density diagnosis
```{r}
pars_Error <- c("Ve_Aurine(1)",	"Ve_CBldU(1)",	"Ve_CKU(1)",	"Ve_CLU(1)",	"Ve_CBrnU(1)")
            
pars_Pop <- c("M_lnPLC(1)", "M_lnPKC(1)", "M_lnPBrnC(1)",	"M_lnPRestC(1)",	"M_lnKabsC(1)",	"M_lnKunabsC(1)",	"M_lnKbileC(1)",	"M_lnKurineC(1)",	"M_lnKbrnC(1)")

pars_Study <- c("lnPLC(1.1)", "lnPKC(1.1)", "lnPBrnC(1.1)",	"lnPRestC(1.1)",	"lnKabsC(1.1)",	"lnKunabsC(1.1)",	"lnKbileC(1.1)",	"lnKurineC(1.1)",	"lnKbrnC(1.1)", "lnPLC(1.2)", "lnPKC(1.2)", "lnPBrnC(1.2)",	"lnPRestC(1.2)",	"lnKabsC(1.2)",	"lnKunabsC(1.2)",	"lnKbileC(1.2)",	"lnKurineC(1.2)",	"lnKbrnC(1.2)", "lnPLC(1.3)", "lnPKC(1.3)", "lnPBrnC(1.3)",	"lnPRestC(1.3)",	"lnKabsC(1.3)",	"lnKunabsC(1.3)",	"lnKbileC(1.3)",	"lnKurineC(1.3)",	"lnKbrnC(1.3)", "lnPLC(1.4)", "lnPKC(1.4)", "lnPBrnC(1.4)",	"lnPRestC(1.4)",	"lnKabsC(1.4)",	"lnKunabsC(1.4)",	"lnKbileC(1.4)",	"lnKurineC(1.4)",	"lnKbrnC(1.4)")

pars_Dist <- c("LnPrior",	"LnPosterior",	"LnData")

stan_report_error <- rstan::monitor(x[,,pars_Error], warmup = 0, digits = 4)
stan_report_pop <- rstan::monitor(x[,,pars_Pop], warmup = 0, digits = 4)
stan_report_Study <- rstan::monitor(x[,,pars_Study], warmup = 0, digits = 4)
stan_report_Dist <- rstan::monitor(x[,,pars_Study], warmup = 0, digits = 4)
```

```{r}
bayesplot::mcmc_dens_overlay(x[,,],  pars = pars_Error)
bayesplot::mcmc_dens_overlay(x[,,],  pars = pars_Pop)
bayesplot::mcmc_dens_overlay(x[,,],  pars = pars_Study)
bayesplot::mcmc_dens_overlay(x[,,],  pars = pars_Dist)

bayesplot::mcmc_trace(x[,,], pars = pars_Error)
bayesplot::mcmc_trace(x[,,], pars = pars_Pop)
bayesplot::mcmc_trace(x[,,], pars = pars_Study)
bayesplot::mcmc_trace(x[,,], pars = pars_Dist)
```

# Andrew Gelman Scale Reduction Factor
```{r}
monitor(x, digits = 4)
```

# MPSRF
```{r}
pars_X <- x[j, , pars_Pop[5]] |> as.data.frame() |> lapply(mcmc) |> as.mcmc.list()
gelman.diag(pars_X)
```

# Andrew Gelman Scale Reduction Factor
```{r}
#micemonitor <-monitor(x, digits=4, probs = c(0.025, 0.5, 0.975), warmup = floor(dim(x)[1]/2))
#print(micemonitor)
#install.packages("writexl")
#library(writexl)
#write_xlsx(x = micemonitor, path = "C:/Users/ysl02/mcsim-6.2.0/IRIS_iHg/modeling/miceHier/miceIHg.xlsx", col_names = TRUE)
```

# Validation
```{r}
pars_name <- dimnames(x)[[3]]
str <- which(pars_name == "M_lnPLC(1)")
end <- which(pars_name == "V_lnKbrnC(1)")
pars <- pars_name[str:end]
mcmc_out <- x[j, , pars]
l <- dim(mcmc_out)[1] * dim(mcmc_out)[2]
dim(mcmc_out) <- c(l, length(pars))
colnames(mcmc_out) <- pars
sample_no <- 50
i <- sample(dim(mcmc_out)[1], sample_no) # random select 10 draws
tmp.x <- mcmc_out[i, pars] 
tmp.x |> write.table(file="poppred.dat", row.names=T, sep="\t")
model <- "iHgMiceBW.model"
input <- "iHgMice_MCMC_setpts.in" 
mcsim(model = model, input = input, dir = "modeling")
```
```{r}
pred_pop <- fread("poppred.out") |> as.data.frame()
dim(pred_pop)
names(pred_pop)
```

```{r}
validation_data <- c(
  data_c_Ntp1993Female1250liver,
  data_c_Ntp1993Female1250kidney,
  data_c_Ntp1993Female5000liver,
  data_c_Ntp1993Female5000kidney,
  data_c_Ntp1993Female20000liver,
  data_c_Ntp1993Female20000kidney,
  data_c_Ntp1993Female20000brain,
  data_c_Ntp1993Male29550liver,
  data_c_Ntp1993Male29550kidney,
  data_c_Ntp1993Male29550brain,
  data_c_Ntp1993Female29550liver,
  data_c_Ntp1993Female29550kidney,
  data_c_Ntp1993Female29550brain,
  data_c_Sin6000liver,
  data_c_Sin6000kidney,
  data_c_Sin6000brain,
  data_c_Wang10000liver,
  data_c_Wang10000kidney,
  data_c_Wang10000brain,
  data_c_Wang10000blood
)
```

```{r}
str <- which(names(pred_pop)=="CLU_1.1")
pred_pop_data <- pred_pop[,c(str:dim(pred_pop)[2])]
organs <- c("Liver", "Liver", "Liver", 
            "Kidney", "Kidney", "Kidney", 
            "Liver", "Liver", "Liver", 
            "Kidney", "Kidney", "Kidney", 
            "Liver", "Liver", "Liver", 
            "Kidney", "Kidney", "Kidney", 
            "Brain", "Brain", "Brain",
            "Liver", "Kidney", "Brain",
            "Liver", "Kidney", "Brain",
            "Liver", "Kidney", "Brain", # Sin
            "Liver", "Kidney", "Brain", "Blood") # Wang
study <- c(rep("Ntp 1993", 27), 
           rep("Sin 1990", 3),
           rep("Wang 2013", 4))
dose <- c(rep(c("925 ug/kg", "3695 ug/kg", "14775 ug/kg"), 6),
          rep("14775 ug/kg", 3),
          rep("29550 ug/kg", 6),
          rep("6000 ug/kg", 3),
          rep("25560 ug/kg", 4))

# check data length
length(na.omit(validation_data))==dim(pred_pop_data)[2]

reshape2::melt(pred_pop_data) |> mutate(obs = rep(na.omit(validation_data), each=sample_no)) |>
  mutate(ratio = value / obs) |> select(variable, ratio) |> 
  mutate(organs = rep(organs, each=sample_no), study= rep(study, each=sample_no), dose=rep(dose, each=sample_no)) |> 
  unite("label", study:dose) |>
  mutate(acceptance = ifelse(ratio > 3, 0, ifelse(ratio < 0.33, 0, 1))) |>
  group_by(label) |>
  mutate(accept_rate = sum(acceptance)/n()) |>
  ggplot() + 
  geom_point(aes(x=ratio, y=label)) +
  scale_x_log10() +
  facet_grid(~organs) +
  geom_vline(xintercept = 1, lty=2) +
  labs(x="Prediction / Observation", y="Study") +
  theme(strip.background = element_blank())
```

```{r}
validation_mice <- reshape2::melt(pred_pop_data) |> mutate(obs = rep(na.omit(validation_data), each=sample_no)) |>
  mutate(ratio = value / obs) |> select(variable, ratio) |> 
  mutate(organs = rep(organs, each=sample_no), study= rep(study, each=sample_no), dose=rep(dose, each=sample_no)) |> 
  unite("label", study:dose) |>
  mutate(acceptance = ifelse(ratio > 3, 0, ifelse(ratio < 0.33, 0, 1))) |>
  mutate(species = "Mice") |> 
  group_by(species, organs) |>
  mutate(accept_rate = sum(acceptance)/n()) 
validation_mice |> distinct(species, organs, accept_rate) 
save(validation_mice, file = "validation_mice.RData")
```

# Software package
# Check the software package version that will be used in this 
# example.
```{r, eval=T}
devtools::session_info()
```

```{r, include=FALSE}
clear()
```
