---
title: "iHg - Markov chain Monte Carlo Calibration"
author: "Nan-Hung Hsieh" modified by "Yu-Sheng Lin"
date: "2023/02/05 (update: `r Sys.Date()`)"
output: 
  html_document:
  fig_caption: yes  
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#setwd('..')
#wd <- getwd()
#knitr::opts_knit$set(root.dir =  wd)
```

# 0 Prerequisites
```{r, include=FALSE}
source("MCSim/function.R")
getwd()
```

# Import the R packages
```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(bayesplot)
library(corrplot)
library(sensitivity)
library(pksensi)
library(data.table)
library(coda)
theme_set(theme_bw())
```
##### 2. QA/QC check


##### 2.1 Balance check and gavage calculation check --- dynamic body weight change only for humans
```{r}
RMCSim::makemcsim("iHgHumanBW.model", dir = "modeling")
test <- mcsim(model = "iHgHumanBW.model", input = "iHgHumanFarrisBW.in", dir = "modeling")
testex <- readsims(test,  exp=1) 
plot (testex$Time, testex$BalanceCheck, xlim = c(0, 2200), ylim = c(-1, 1),  type="l", lwd=2, xlab="Time", ylab="mass balance")
```

###### 3.1 Model verification of Individual studies ########

#################### human study #################

##### 1. Human volunteers (assumed BW of 70 kg)
# Single IV dose of 0.025 µg Hg/kg [0.6–2.8 µg Hg from 
# 203Hg(NO3)2] (Farris, 2008)
```{r}
Farris <- mcsim(model = "iHgHumanBW.model", input = "iHgHumanFarrisBW.in", dir = "modeling")
Farris <- readsims(Farris, exp=1) 
```

##### 1. Human w/ single IV dose of 0.025 µg Hg/kg  (Farris, 2008): feces    
```{r}
data_t <- c(24 , 48 , 72 , 96 , 120 , 144 , 168 , 336 , 504 , 672 , 840 , 1008 , 1176 , 1344 , 1512 , 1680);
data_c <- c(0.05 , 0.08148 , 0.09493671 , 0.10810127 , 0.15063291 , 0.15949367 , 0.1733971 , 0.21359439 , 0.2432366 , 0.27616296 , 0.29439456 , 0.32346864 , 0.34659148 , 0.36122036 , 0.38278073 , 0.40246241);
      
plot(data_t, data_c, main = "Human Farris, 2008, IV=0.025", xlab = "Time (hr)", ylab = "Afeces (ug)", xlim = c(0, 1700), ylim = c(0.01, 10), log='y')
lines(Farris$Time, Farris$Afeces)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 1. Human w/ single IV dose of 0.02 µg Hg/kg  (Farris, 2008): urine    
```{r}
data_t <- c(24 , 48 , 72 , 96 , 120 , 144 , 168 , 336 , 504 , 672 , 840 , 1008 , 1176 , 1344 , 1512 , 1680);
data_c <- c(0.01016 , 0.02318471 , 0.02675159 , 0.03388535 , 0.03923567 , 0.04458599 , 0.04902761 , 0.0929182 , 0.12125111 , 0.16541369 , 0.19404394 , 0.23234815 , 0.26564472 , 0.29066447 , 0.31457481 , 0.34118414);

plot(data_t, data_c, main = "Human Farris, 2008, IV=0.02", xlab = "Time (hr)", ylab = "Aurine (ug)", xlim = c(0, 1700), ylim = c(0.01, 1), log='y')
lines(Farris$Time, Farris$Aurine)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 1. Human w/ single IV dose of 0.025 µg Hg/kg  (Farris, 2008): blood in the unit of ug/ml   
```{r}
data_t <- c(24 , 72 , 96 , 168 , 336 , 504 , 672 , 840 , 1008 , 1176 , 1344 , 1512, 1680);
data_c <- c(0.0972 , 0.0377 , 0.0257 , 0.0171 , 0.0103 , 0.0082 , 0.0069 , 0.0061 , 0.0062 , 0.0051 , 0.0045 , 0.0046 ,0.0039);

plot(data_t, data_c, main = "Human Farris, 2008, IV=0.02", xlab = "Time (hr)", ylab = "ABld (ug)", xlim = c(0, 1700), ylim = c(0.001, 10), log='y')
lines(Farris$Time, Farris$ABld)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

# conc. data_c <- c(0.981605057,		0.388320929	,	0.175181914	,	0.105565314	,	0.084128514	,	0.070550414	,	# 0.062251586	# ,	0.064024129	,	0.052330886	,	0.046171486	,	0.047486971,	0.039810071);



##### 2. Human volunteers (BW = 64 kg) (Rahola, 1973) 
# Single oral dose of 0.09375 µg Hg/kg (6 µg Hg from 203Hg(NO3)2]j 
```{r}
Rahola <- mcsim(model = "iHgHumanBW.model", input = "iHgHumanRaholaBW.in", dir = "modeling")
Rahola <- readsims(Rahola, exp=1) 
```

##### 2. Human w/ single oral dose of 0.09375 µg Hg/kg  (Rahola, 1973): feces    
```{r}
data_t <- c(24 , 48 , 72 , 96 , 144 , 192 , 240 , 288 , 336 , 384 , 408 , 432 , 456 , 528 , 672 , 720 , 1224 , 1248 , 1272);
data_c <- c(1.477809 , 3.470468 , 4.724273 , 5.272337 , 5.665805 , 5.774183 , 5.811173 , 5.826059 , 5.835748 , 5.842266 , 5.847794 , 5.855498 , 5.859217 , 5.863198 , 5.864334 , 5.865867 , 5.866668 , 5.867613 , 5.868693);

plot(data_t, data_c, main = "Human Rahola, 1973, oral=0.09375", xlab = "Time (hr)", ylab = "Afeces (ug)", xlim = c(0, 1300), ylim = c(0.1,100), log='y')
lines(Rahola$Time, Rahola$Afeces)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 2. Human w/ single oral dose of 0.09375 µg Hg/kg  (Rahola, 1973): urine    
```{r}
data_t <- c(24 , 48 , 72 , 96 , 144 , 192 , 240 , 288 , 336 , 384 , 408 , 432 , 456 , 528 , 672 , 720 , 1224 , 1248 , 1272 , 1824 , 1848);
data_c <- c(0.001754597 , 0.004511698 , 0.006563485 , 0.008784912 , 0.012570932 , 0.016304637 , 0.020405739 , 0.022484365 , 0.028068776 , 0.030784725 , 0.035378559 , 0.042058603 , 0.045490943 , 0.050140927 , 0.056225823 , 0.063441201 , 0.064927833 , 0.067125743 , 0.068868326 , 0.076238348 , 0.076842953);

plot(data_t, data_c, main = "Human Rahola, 1973, oral=0.09375", xlab = "Time (hr)", ylab = "Aurine (ug)", xlim = c(0, 1850), ylim = c(0.001, 1), log='y')
lines(Rahola$Time, Rahola$Aurine)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```


##### 3. Human volunteers (BW = 48 kg) (Yoshida, 1997) 
# Single oral dose of Hg (0.66g)
```{r}
Yoshida <- mcsim(model = "iHgHumanBW.model", input = "iHgHumanYoshidaBW.in", dir = "modeling")
Yoshida <- readsims(Yoshida, exp=1) 
```


##### 3. Human w/ single oral dose of 13750 µg Hg/kg  (Yoshida, 1997): urine 
```{r}
data_t <- c(528 , 562 , 600 , 715 , 763 , 811 , 859 , 907 , 955 , 1022 , 1070 , 1094 , 1118 , 1142 , 1166 , 1186 , 1243 , 1262 , 1286 , 1301 , 1330 , 1382 , 1435 , 1450 , 1474 , 1550 , 1613 , 1642 , 1666 , 1680 , 1742 , 1786 , 1810 , 1834 , 1858 , 1901 , 1973 , 2002 , 2030 , 2069 , 2213 , 2242);

data_c <- c(409.0 , 934.6 , 1603.7 , 2516.1 , 3439.6 , 4446.1 , 5015.7 , 5585.3 , 6204.6 , 6707.8 , 7365.9 , 7841.5 , 8449.8 , 9107.8 , 9644.2 , 10263.6 , 11065.4 , 11568.7 , 12530.9 , 12934.6 , 13343.8 , 13813.8 , 14234.1 , 14632.3 , 15074.7 , 15257.1 , 15561.3 , 15826.7 , 16119.8 , 16556.7 , 16827.6 , 17087.6 , 17424.9 , 17612.9 , 17812.0 , 17972.4 , 18177.0 , 18436.9 , 18785.3 , 19034.1 , 19249.8 , 19581.6);

data_c_Yoshida1997urine <- data_c

plot(data_t, data_c, main = "Human Yoshida, 1997 Single oral of 0.66 g", xlab = "Time (hr)", ylab = "Aurine (ug)", xlim = c(0, 1850), ylim = c(1, 30000), log='y')
lines(Yoshida$Time, Yoshida$Aurine)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 3. Human w/ single oral of 13750 µg Hg/kg  (Yoshida, 1997): blood  
```{r}
data_t <- c(315	,	1213	,	1328	,	1500	,	1672	,	1835	,	2011	,	2174);
data_c <- c(1.410	,	0.728	,	0.543	,	0.449	,	0.250	,	0.253	,	0.204	,	0.166);

plot(data_t, data_c, main = "Human Yoshida, 1997, Single oral of 0.66 g", xlab = "Time (hr)", ylab = "CBldU (ug/ml)", xlim = c(0, 3300), ylim = c(0.01, 10), log='y')
lines(Yoshida$Time, Yoshida$CBldU)
legend("topright", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

##### 4. Monkey (BW = 3.53 kg) (Vahter, 1994) 
# IV (608 ug/day or ug/kg/day)
```{r}
Vahter <- mcsim(model = "iHgHumanBW.model", input = "iHgMonkeyVahterBW.in", dir = "modeling")
Vahter <- readsims(Vahter, exp=1) 
```

# 4.  Monkey (BW = 3.53 kg) (Vahter, 1994) : brain
```{r}
data_t <- c(1272);
data_c <- c(0.09317);
sd <- c(0.03)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Vahter, 1994, daily IV, Monkey", xlab = "Time (hr)", ylab = "CBrn (ug/g)", xlim = c(0, 1300), ylim = c(0.01, 10), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Vahter$Time, Vahter$CBrnU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

# 4.  Monkey (BW = 3.53 kg) (Vahter, 1994) : blood
```{r}
data_t <- c(168);
data_c <- c(0.6);
sd <- c(0.14)
sd.up <- data_c + sd
sd.dn <- data_c - sd

plot(data_t, data_c, main = "Vahter, 1994, daily IV, Monkey", xlab = "Time (hr)", ylab = "CBld (ug/g)", xlim = c(0, 200), ylim = c(0.01, 10), log='y')
arrows(data_t, sd.dn, data_t, sd.up, code=3, length=0.02, angle=90)
lines(Vahter$Time, Vahter$CBldU)
legend("topleft", legend = c("Data", "Model"), lty = c(NA, 1), pch = c(1, NA))
```

# Single chain check
# Checking model fit using one single chain #
# Global evaluation/ "MCMC.check.out" has "Data" and "Predicion"
```{r}
single_1 <- fread("outputs/iHgHuman_4880.out") |> as.data.frame()
tail(single_1)
dim(single_1)
```

## Result diagnosis ##
# Trace plot
```{r}
str <- ceiling(nrow(single_1)/2) + 1
end <- nrow(single_1)
j <- c(str:end) # discard burn-in
par(mfrow = c(3,4), mar = c(2,2,2,1))
for (i in 2:dim(single_1)[2]){
  plot(single_1[j,1], single_1[j,i], type = "l", main = names(single_1)[i])
}
```

# Auto correlation
```{r}
par(mfrow = c(3,4), mar = c(2,2,4,1))
for (i in 2:dim(single_1)[2]){
  acf(single_1[j,i], main = names(single_1)[i])
}
```

# Density plot
```{r}
par(mfrow = c(3,4), mar = c(2,2,2,1))
for (i in 2:dim(single_1)[2]){
  single_1[j,i] %>% density() %>% plot(type = "l", main = names(single_1)[i])
}
```

# Correlation matrix
```{r}
single_1[,4:12] %>% cor() %>% corrplot(method = "number")
```

# Global evaluation/ "MCMC.check.out" has "Data" and "Predicion"
# Check simulation result with the parameter from the last iteration
```{r}
Humanchk_1 <- read.delim("outputs/iHgHuman_check_4880.out")
Humanchk_1 <- Humanchk_1[which(Humanchk_1$Data>0),]
tail(Humanchk_1)
```

# Overall fitting evaluation
```{r}
R_square <- cor(Humanchk_1$Data, Humanchk_1$Prediction)^2
plot(Humanchk_1$Data, Humanchk_1$Prediction, log = "xy",
     xlim = c(0.001, 1000), ylim = c(0.001, 1000),
     xlab = "Observation for iHg in humans", ylab = "Prediction for iHg in humans",
     main = paste("R2 =", round(R_square, digits = 3)))
abline(0,1)
```

# Individual measurement fitting evaluation 
# (1) amount of iHg in feces
```{r}
Humanchk_1 %>% filter(Output_Var == "Afeces") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "amount (ug)", title = "Accumulative amount of excreted iHg in feces (ug) in humans")
```

# (2) amount of iHg in urine
```{r}
Humanchk_1 %>% filter(Output_Var == "Aurine") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "amount (ug)", title = "Accumulative amount of excreted iHg in urine (ug) in humans")
```

#### 2.5 Multi chains testing ### 
#### MCMC Modeling ###
### Using Ubuntu instead ###

```{r, eval=F}
Humanjob1 <- fread("outputs/iHgHuman_4880.out") |> data.frame()
Humanjob2 <- fread("outputs/iHgHuman_3365.out") |> data.frame()
Humanjob3 <- fread("outputs/iHgHuman_5916.out") |> data.frame()
Humanjob4 <- fread("outputs/iHgHuman_6734.out") |> data.frame()
x <- mcmc_array(data = list(Humanjob1, Humanjob2, Humanjob3, Humanjob4))
str <- ceiling(nrow(x)/2) + 1
end <- nrow(x)
j <- c(str:end) # discard burn-in
```

# Size of output summary
```{r}
dim(x)
# X [iteration, chain, parameter]
```
# traceplot and density diagnosis
```{r}
pars_Error <- c("Ve_Aurine.1.",	"Ve_Afeces.1.",	"Ve_CBldU.1.",	"Ve_CKU.1.",	"Ve_CLU.1.",	"Ve_CBrnU.1.")
            
pars_Pop <- c("M_lnPLC.1.", "M_lnPKC.1.", "M_lnPBrnC.1.",	"M_lnPRestC.1.",	"M_lnKabsC.1.",	"M_lnKunabsC.1.",	"M_lnKbileC.1.",	"M_lnKurineC.1.",	"M_lnKbrnC.1.")

pars_Study <- c("lnPLC.1.1.", "lnPKC.1.1.", "lnPBrnC.1.1.",	"lnPRestC.1.1.",	"lnKabsC.1.1.",	"lnKunabsC.1.1.",	"lnKbileC.1.1.",	"lnKurineC.1.1.",	"lnKbrnC.1.1.", "lnPLC.1.2.", "lnPKC.1.2.", "lnPBrnC.1.2.",	"lnPRestC.1.2.",	"lnKabsC.1.2.",	"lnKunabsC.1.2.",	"lnKbileC.1.2.",	"lnKurineC.1.2.",	"lnKbrnC.1.2.")

pars_Dist <- c("LnPrior",	"LnPosterior",	"LnData")

mcmc_dens_overlay(x[j,,],  pars = pars_Error)
mcmc_dens_overlay(x[j,,],  pars = pars_Pop)
mcmc_dens_overlay(x[j,,],  pars = pars_Study)
mcmc_dens_overlay(x[j,,],  pars = pars_Dist)

mcmc_trace(x[j,,], pars = pars_Error)
mcmc_trace(x[j,,], pars = pars_Pop)
mcmc_trace(x[j,,], pars = pars_Study)
mcmc_trace(x[j,,], pars = pars_Dist)
```

# Andrew Gelman Scale Reduction Factor
```{r}
monitor(x, digits = 4)
```


# Validation
```{r}
pars_name <- dimnames(x)[[3]]
str <- which(pars_name == "M_lnPLC.1.")
end <- which(pars_name == "V_lnKbrnC.1.")
pars <- pars_name[str:end]
mcmc_out <- x[j, , pars]
l <- dim(mcmc_out)[1] * dim(mcmc_out)[2]
dim(mcmc_out) <- c(l, length(pars))
colnames(mcmc_out) <- pars
i <- sample(dim(mcmc_out)[1], 10) # random select 10 draws
tmp.x <- mcmc_out[i, pars] 
tmp.x |> write.table(file="poppred.dat", row.names=T, sep="\t")
model <- "iHgHumanBW.model"
input <- "iHgHuman_MCMC_setpts.in" 
mcsim(model = model, input = input, dir = "modeling")
```

```{r}
pred_pop <- fread("poppred.out") |> as.data.frame()
dim(pred_pop)
names(pred_pop)
```

```{r}
validation_data <- data_c_Yoshida1997urine |> tail(1)
pred_pop_data <- pred_pop[,dim(pred_pop)[2]]
organs <- "Urine"
study <- "Yoshida"
dose <- "13750 ug/kg"

# check data length
reshape2::melt(pred_pop_data) |> mutate(obs = rep(na.omit(validation_data), each=10)) |>
  mutate(ratio = value / obs) |> select(ratio) |> 
  mutate(organs = rep(organs, 10), study= rep(study, 10), dose=rep(dose, 10)) |> 
  unite("label", study:dose) |>
  ggplot() + 
  geom_point(aes(x=ratio, y=label)) +
  scale_x_log10() +
  facet_grid(~organs) +
  geom_vline(xintercept = 1, lty=2) +
  labs(x="Prediction / Observation", y="Study") +
  theme(strip.background = element_blank())
```

```{r}
validation_human <- reshape2::melt(pred_pop_data) |> mutate(obs = rep(na.omit(validation_data), each=10)) |>
  mutate(ratio = value / obs) |> select(ratio) |> 
  mutate(organs = rep(organs, 10), study= rep(study, 10), dose=rep(dose, 10)) |> 
  unite("label", study:dose) |> mutate(species = "Human")
save(validation_human, file = "validation_human.RData")
```

# Software package
# Check the software package version that will be used in this 
# example.
```{r, eval=T}
devtools::session_info()
```

```{r, include=FALSE}
clear()
```
